Type Refinements for the Working Class

### 2015-11-14 paper.tex
Begin explaining lazy computation systems

\newcommand\ArrId[1]{1_{#1}}
\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}

\newcommand\Symbols{\mathbf{I}}
\newcommand\SymSets{\mathbb{F}_\Symbols}
\newcommand\FinSets{\mathbb{F}}
\newcommand\Sorts{\mathcal{S}}
\newcommand\Valences{\mathcal{V}}
\newcommand\Arities{\mathcal{A}}
\newcommand\Operators{\mathcal{O}}

@@

\newcommand\IsEq[2]{\IMode{#1} = \IMode{#2}}
\newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}

@@

@@

>>>>

\newcommand\ArrId[1]{1_{#1}}
\newcommand\Match[2]{\IMode{#1}\equiv\OMode{#2}}
\newcommand\MapsTo[2]{\IMode{#1}\mapsto\OMode{#2}}

\newcommand\Symbols{\mathbf{I}}
\newcommand\SymSets{\mathbb{F}_\Symbols}
\newcommand\FinSets{\mathbb{F}}
\newcommand\Sorts{\mathcal{S}}
\newcommand\Valences{\mathcal{V}}
\newcommand\Arities{\mathcal{A}}
\newcommand\Operators{\mathcal{O}}
\newcommand\CanOperators{\mathcal{K}}

@@

\newcommand\IsEq[2]{\IMode{#1} = \IMode{#2}}
\newcommand\IsApart[2]{\IMode{#1} \mathrel{\#} \IMode{#2}}
\newcommand\IsSubsetEq[2]{\IMode{#1}\subseteq\IMode{#2}}

@@

\newcommand\TyBase[1]{\mathtt{Base}_{#1}}
\newcommand\TyTop[1]{\top_{#1}}

\newcommand\EvalN[5]{\IMode{#1}\parallel\IMode{#4}\Downarrow^{\OMode{#3}}_{\IMode{#2}}\OMode{#5}}
\newcommand\Exprs{\mathcal{E}}
\newcommand\Values{\mathcal{V}}
\newcommand\Naturals{\mathbb{N}}
\newcommand\NaturalsPsh{\mathcal{N}}

@@

\section{Multi-Sorted Computational Type Theory}

Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
you specify how to form its equal canonical inhabitants, or values. Then, the
inhabitants of the types are explained uniformly by extending this relation
over an operational semantics such that two terms are equal members of a
type just when they compute to equal values of that type.

More generally, one may wish to consider types which include computations in
addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
is a type, which would include all the members of $\psi$ as well as a divergent
computation. Another example is the type $\TyTop{\tau}$ which identifies all
closed terms of sort $\tau$.

In order to ensure that such types are sensible, we must formulate a notion of
computational equivalence, and then say that a refinement is a \emph{type} just
when it respects computational equivalence; then, we are able to define types
as PERs on \emph{programs} in addition to values.

\subsection{Lazy Computation Systems}

In this section we generalize Howe's notion of lazy computation
system~\cite{howe:1989} to the multi-sorted, symbol-parameterized setting. An
\emph{lazy computation language} is an abt signature
$\Match{\Sigma}{\Pair{\Sorts}{\Operators}}$ along with a distinguished presheaf
$\Of{\CanOperators}{\Sets^{\SCtx\times\Discrete{\Arities}}}$ of
\emph{canonical} operators such that $\IsSubsetEq{\CanOperators}{\Operators}$.
For a lazy computation language
$\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define the
copresheaves of closed expressions and closed values over symbol contexts as
follows:
\begin{align*}
  \ADefine{\Exprs_\tau(\Upsilon)}{
    \MkSet{M \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{M}{\tau}}
  }\\
  \ADefine{\Values_\tau(\Upsilon)}{
    \MkSet{
      M\equiv\App{\vartheta}{\vec{E}}
      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators\Pair{\Upsilon}{a}\ni\vartheta
    }
  }
\end{align*}

Then, a \emph{lazy computation system} (lcs) is a lazy computation language
$\Match{L}{\Pair{\Sigma}{\CanOperators}}$ along with an indexed
$\alpha$-functional relation
$\Of{\Downarrow}{\Pow{\Exprs\times\Values\times\NaturalsPsh}}$ where
$\NaturalsPsh$ is the constant presheaf $\MapsTo{(\Upsilon, \tau)}{\Naturals}$.
We will write $\EvalN{\Upsilon}{\tau}{n}{M}{N}$ when
$\Member{\Tuple{M,N,n}}{\Downarrow_\tau(\Upsilon)}$, to mean that $M$ evaluates
to $N$ in $n$ steps.
